Follow-up Samples

### 2017-05-14 submission/manuscript.tex
fix superscript and add actual ranges

Marc A Sze${^1}$, Nielson T Baxter${^1,2}$, Mack T Ruffin IV${^3}$, Mary AM Rogers${^2}$, and Patrick D Schloss${^1}$${^\dagger}$

@@

There was a subset of patients (6 of the 26 with carcinomas) who
demonstrated an elevated probability of carcinoma after treatment. This
may reflect an elevated risk of recurrence. The 23.08\% prevalence of
increased carcinoma probability from our study is within the expected
rate of recurrence (20-30\% {[}3,4{]}). We hypothesized that these
individuals may have had more severe tumors; however, the tumor severity
of these 6 individuals (3 with Stage II and 3 with Stage III) was
similar to the distribution observed among the other 20 patients. We
also hypothesized that we may have sampled these patients later than the
rest and their communities may have reverted to a carcinoma-associated
state; however, there was not a statistically significant difference in
the length of time between sample collection among those whose
probabilities increased (358 (121) days) or decreased (334 (143) days)
(Wilcoxon Test; P-value = 0.56) (all days data displayed as median
(IQR)). Finally, it is possible that these patients may not have
responded to treatment as well as the other 20 patients diagnosed with
carcinoma and so the microbiota may not have been impacted the same way.
Again, further studies looking at the role of the microbiota in
recurrence are needed to understand the dynamics following treatment.

Our final hypothesis was that the specific type of treatment altered the
structure of the microbiome. The treatment to remove adenomas and
advanced adenomas was either polyp removal or surgical resection whereas
it was surgical resection alone or in combination with chemotherapy or
with chemotherapy and radiation for individuals with carcinoma. Because
chemotherapy and radiation target rapidly growing cells, these
treatments would be more likely to cause a turnover of the colonic
epithelium driving a more significant change in the structure of the
microbiota. Although, we were able to test for an effect across these
specific types of treatment, the number of patients in each treatment
group was relatively small. Finally, those undergoing surgery would have
received antibiotics and this may be a potential confounder. However,
our pre-treatment stool samples were obtained before the surgery and the
post-treatment samples were obtained long after any effects due to
antibiotic administration on the microbiome would be expected to occur
(344 (142) days).

>>>>

Marc A Sze${^1}$, Nielson T Baxter${^1}$${^,}$${^2}$, Mack T Ruffin IV${^3}$, Mary AM Rogers${^2}$, and Patrick D Schloss${^1}$${^\dagger}$

@@

There was a subset of patients (6 of the 26 with carcinomas) who
demonstrated an elevated probability of carcinoma after treatment. This
may reflect an elevated risk of recurrence. The 23.08\% prevalence of
increased carcinoma probability from our study is within the expected
rate of recurrence (20-30\% {[}3,4{]}). We hypothesized that these
individuals may have had more severe tumors; however, the tumor severity
of these 6 individuals (3 with Stage II and 3 with Stage III) was
similar to the distribution observed among the other 20 patients. We
also hypothesized that we may have sampled these patients later than the
rest and their communities may have reverted to a carcinoma-associated
state; however, there was not a statistically significant difference in
the length of time between sample collection among those whose
probabilities increased (358 (336 - 458) days) or decreased (334 (256 -
399) days) (Wilcoxon Test; P-value = 0.56) (all days data displayed as
median (IQR)). Finally, it is possible that these patients may not have
responded to treatment as well as the other 20 patients diagnosed with
carcinoma and so the microbiota may not have been impacted the same way.
Again, further studies looking at the role of the microbiota in
recurrence are needed to understand the dynamics following treatment.

Our final hypothesis was that the specific type of treatment altered the
structure of the microbiome. The treatment to remove adenomas and
advanced adenomas was either polyp removal or surgical resection whereas
it was surgical resection alone or in combination with chemotherapy or
with chemotherapy and radiation for individuals with carcinoma. Because
chemotherapy and radiation target rapidly growing cells, these
treatments would be more likely to cause a turnover of the colonic
epithelium driving a more significant change in the structure of the
microbiota. Although, we were able to test for an effect across these
specific types of treatment, the number of patients in each treatment
group was relatively small. Finally, those undergoing surgery would have
received antibiotics and this may be a potential confounder. However,
our pre-treatment stool samples were obtained before the surgery and the
post-treatment samples were obtained long after any effects due to
antibiotic administration on the microbiome would be expected to occur
(344 (266 - 408) days).
